---
title: "class09"
author: "Jaewon Kim"
format: pdf
---

```{r}
candy_files <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

candy = read.csv(candy_files, row.names = 1)
head(candy)
```

Q1. How many different candy types are in this dataset?
```{r}
nrow(candy)
```
Row is candies, while column is criteria. Therefore, there are 85 types of candies. 


Q2. How many fruity candy types are in the dataset?
```{r}
sum(candy$fruity)
```
Candies that are fruity has 1 in cell. Therefore, there are 38 fruity candies.


Q3. What is your favorite candy in the dataset and what is it’s winpercent value?
```{r}
candy["Skittles original", ]$winpercent
```
My favorite candy is Skittles, and its winpercent is 63.085%.


Q4. What is the winpercent value for “Kit Kat”?
```{r}
candy["Kit Kat", ]$winpercent
```
Winpercent of Kit Kat is 76.769%.


Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?
```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```
Winpercent of Tootsie Roll Snack Bars is 49.654%. 


```{r}
#install.packages("skimr")
library("skimr")
skim(candy)
```

Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
- p0 to p100 of chocolate to pluribus is either zero or one. However, column values for sugarpercent to winpercent are in continuous numeric value between zero to one or zero to hundred (because they're percent data). Therefore, variable chocolate~pluribus/sugarpercent~winpercent have different scale for column p0~p100.


Q7. What do you think a zero and one represent for the candy$chocolate column?
- Zero means that corresponding candies doesn't contain chocolate.


Q8. Plot a histogram of winpercent values
```{r}
library(ggplot2)
library(dplyr)

ggplot(candy, aes(x = candy$winpercent)) + 
  geom_histogram() +
  xlab("winpercent (%)") +
  ylab("count") +
  ggtitle("winpercent of candies")
```

Q9. Is the distribution of winpercent values symmetrical?
- No, distribution shows right skewness.


Q10. Is the center of the distribution above or below 50%?
- Since distribution is positive skew, center of the distribution is below 50%.


Q11. On average is chocolate candy higher or lower ranked than fruit candy?
```{r}
choc <- candy$winpercent[as.logical(candy$chocolate)]
can <- candy$winpercent[as.logical(candy$fruity)]
mean(choc)
mean(can)
```
Average winpercent of chocolate candy is 60.922%, while fruity candy is 44.120%. Therefore, chocolate candy is ranked higher than fruit candy. 


Q12. Is this difference statistically significant?
```{r}
t.test(choc, can)
```
Calculated t value was 6.25, and corresponding p-value at 68.882 degree of freedom was less than 0.001. Low p-value fails to reject null hypothesis, hence, there is no statistically significant evidence that tow groups are significantly different. Therefore, no


Q13. What are the five least liked candy types in this set?
```{r}
candy %>%
  arrange(winpercent) %>%
  head(5)
```
From least favorite, five least liked candies are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters


Q14. What are the top 5 all time favorite candy types out of this set?
```{r}
candy %>%
  arrange(desc(winpercent)) %>%
  head(5)
```
From most favorite, five all time favorite candies are Reese's Peanut Butter cup, Reese's Miniatures, Twix, Kit Kat, and Snickers


Q15. Make a first barplot of candy ranking based on winpercent values.
```{r}
ggplot(candy) +
  aes(x = rownames(candy), y = winpercent) +
  geom_col() +
  coord_flip() +
  theme(text = element_text(size = 6))
```


Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?
```{r}
p<- ggplot(candy) +
  aes(x = reorder(rownames(candy),winpercent), y = winpercent) +
  geom_col() +
  coord_flip() +
  theme(text = element_text(size = 6))
p
```


```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

p +
  geom_col(fill = my_cols)
```

Now, for the first time, using this plot we can answer questions like:
- Q17. What is the worst ranked chocolate candy?
Lowest bar with chocolate color is Sixlets. Hence, sixlets is the worst ranked chocolate candy.


- Q18. What is the best ranked fruity candy?
Highest bar with pink color is Starburst. Hence, Starburst is the best ranked fruity candy.


plot privepercent vs. winpercent
```{r}
#install.packages("ggrepel")
library(ggrepel)

ggplot(candy) +
  aes(x = winpercent, y = pricepercent, label = rownames(candy)) +
  geom_point(col = my_cols) +
  geom_text_repel(col = my_cols, size = 3.3, max.overlaps = 5)
```


Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

pricepercent shows percentile rank of unit price, and winpercent shows percentage of wins. Therefore, highest winpercent with least money (lowest pricepercent) is Reese's Miniatures.


Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
Five most expensive candies are Nik L Nip, Nestle Smarties, Ring pop, Hershey's Krackel, and Hershey's Milk Chocolate. Least popular among them is Nik L Nip.


Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().
```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
  geom_point() +
  theme(text = element_text(size = 6))
```


```{r}
#install.packages("corrplot")
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Fruity candy and Chocolate candies are most anti-correlated.


Q23. Similarly, what two variables are most positively correlated?

Besides diagonal, Chocolate and bar, winpercent and chocolate are most positively correlated. 


```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)

plot(pca$x[ ,1], pca$x[ ,2])
plot(pca$x[ ,1:2], col=my_cols, pch=16)
```
```{r}
my_data <- cbind(candy, pca$x[,1:3])

pp <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

pp
```
```{r}
pp + 
  geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space", subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)", caption="Data from 538")
```
```{r}
#install.packages("plotly")
library(plotly)

ggplotly(pp)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard, and pluribus are strongly affecting PC1 into positive direction. This result make sense because all three variables are either zero or one, where candies must align to trend or have huge deviation (when each variables are plotted).



